
---
title: "Mini Data Analysis Milestone 2"
output: github_document
---

*To complete this milestone, you can either edit [this `.rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-2.Rmd) directly. Fill in the sections that are commented out with `<!--- start your work here--->`. When you are done, make sure to knit to an `.md` file by changing the output in the YAML header to `github_document`, before submitting a tagged release on canvas.*

# Welcome to the rest of your mini data analysis project!

In Milestone 1, you explored your data. and came up with research questions. This time, we will finish up our mini data analysis and obtain results for your data by: 

- Making summary tables and graphs 
- Manipulating special data types in R: factors and/or dates and times.
-   Fitting a model object to your data, and extract a result.
-   Reading and writing data as separate files.

We will also explore more in depth the concept of *tidy data.* 

**NOTE**: The main purpose of the mini data analysis is to integrate what you learn in class in an analysis. Although each milestone provides a framework for you to conduct your analysis, it's possible that you might find the instructions too rigid for your data set. If this is the case, you may deviate from the instructions -- just make sure you're demonstrating a wide range of tools and techniques taught in this class.

# Instructions

**To complete this milestone**, edit [this very `.Rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-2.Rmd) directly. Fill in the sections that are tagged with `<!--- start your work here--->`.

**To submit this milestone**, make sure to knit this `.Rmd` file to an `.md` file by changing the YAML output settings from `output: html_document` to `output: github_document`. Commit and push all of your work to your mini-analysis GitHub repository, and tag a release on GitHub. Then, submit a link to your tagged release on canvas.

**Points**: This milestone is worth 50 points: 45 for your analysis, and 5 for overall reproducibility, cleanliness, and coherence of the Github submission. 

**Research Questions**: In Milestone 1, you chose two research questions to focus on. Wherever realistic, your work in this milestone should relate to these research questions whenever we ask for justification behind your work. In the case that some tasks in this milestone don't align well with one of your research questions, feel free to discuss your results in the context of a different research question.

# Learning Objectives

By the end of this milestone, you should:

-   Understand what *tidy* data is, and how to create it using `tidyr`.
-   Generate a reproducible and clear report using R Markdown.
-   Manipulating special data types in R: factors and/or dates and times.
-   Fitting a model object to your data, and extract a result.
-   Reading and writing data as separate files.

# Setup

Begin by loading your data and the tidyverse package below:

```{r, message = FALSE}
library(datateachr) # <- might contain the data you picked!
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# Task 1: Process and summarize your data 

From milestone 1, you should have an idea of the basic structure of your dataset (e.g. number of rows and columns, class types, etc.). Here, we will start investigating your data more in-depth using various data manipulation functions. 

### 1.1 (1 point) 

First, write out the 4 research questions you defined in milestone 1 were. This will guide your work through milestone 2:

<!-------------------------- Start your work below ---------------------------->
1. *What is the density of building registered year?*
2. *How are the number of storeys distributed among buildings?*
3. *How is the distribution of separate gas, hydro, and water meter services among buildings?*
4. *What is the relationship between window types and year built?*
<!----------------------------------------------------------------------------->

Here, we will investigate your data using various data manipulation and graphing functions.

### 1.2 (8 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question.

Ensure that the output of each operation is printed!

Also make sure that you're using dplyr and ggplot2 rather than base R. Outside of this project, you may find that you prefer using base R functions for certain tasks, and that's just fine! But part of this project is for you to practice the tools we learned in class, which is dplyr and ggplot2.

**Summarizing:**

1.  Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2.  Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
3.  Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4. Compute the proportion and counts in each category of one categorical variable across the groups of another categorical variable from your data. Do not use the function `table()`!

**Graphing:**

6. Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read.
7. Make a graph where it makes sense to customize the alpha transparency.

Using variables and/or tables you made in one of the "Summarizing" tasks: 

8. Create a graph that has at least two geom layers. 
9. Create 3 histograms, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

<!------------------------- Start your work below ----------------------------->
**Q1 What is the density of building registered year?**

**Preparation**
```{r}
#overview
class(apt_buildings$year_registered)
summary(apt_buildings$year_registered)
summary(apt_buildings$year_built) 

#clean
Q1cleaned <-apt_buildings %>%
  select(id, year_built, year_registered) %>%
  filter(!is.na(year_registered) & !is.na(year_built)) 
head(Q1cleaned)
```

**Q1.1 Summarizing**

*Task*: Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks (Task 3).
numeric variable: year_registered
categorical variable: year_reg_group with 3 groups (year before 2017, year 2017, year after 2017)

*Comment*: This task is helpful. While designing the question at the first stage for MDA1, I thought year_registered would be like year_built and a density plot is applicable. Indeed, the original data set apt_buildings gives year_registered dbl data type. However, through _"explore year_registered"_, I found the range and variety of year_registered is not as wide as year_built. While year_built ranges from 1805-2019, buildings mainly registered in 2017. Thereby, I set a new categorical variable named year_reg_group with 3 groups (year before 2017, year 2017, year after 2017) with an assumption that 2017 is an identical year for building registration and management. This task thus can serve as a preparation for future tasks to explore the City of Toronto's advancement in managing apartment buildings, which is aligned with the primary goal of bringing up _Question 1_.

```{r}
#create a new variable
answer1.1 <-Q1cleaned%>%
  mutate(year_reg_group=factor(case_when(
      year_registered < 2017 ~ "year before 2017",
      year_registered == 2017 ~ "year 2017",
      year_registered > 2017 ~ "year after 2017"),
    levels=c('year before 2017','year 2017','year after 2017')))
head(answer1.1)
```

**Q1.2 Graphing**

*Task*: Make a graph where it makes sense to customize the alpha transparency (Task 7).

*Comment*: This task is not helpful. Alpha transparency makes sense to _answer1.2_, using geom_point. This is because the darker color presents higher density. Although _answer1.2_ is helpful to understand Question 1, I think geom_dotplot is a better option to visualize the data. However, transparency doesn't make sense to this dotplot. This dotplot is better because the length differences in the graph is easier to be distinguished by human eyes than nuances in transparencies. _answer1.2_better_ makes it easier to compare the amount of buildings in their built years within and among registration year groups.
```{r}
#Graph for the Task:
answer1.2 <- answer1.1%>%
  ggplot(aes(year_reg_group, year_built))+
  geom_point(alpha=0.1) + #customize the transparency
  scale_x_discrete(drop = FALSE) + # Don't drop the year before 2017
  labs(title="The Dot Plot of Built Year by Registration Year Cateogory from 1805 to 2020 ", x= "Registration Year Cateogory", y="Built Year")
  
print(answer1.2)

#A better option
answer1.2_better <- answer1.1%>%
  ggplot(aes(year_reg_group, year_built))+
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth=0.7) +
  scale_x_discrete(drop = FALSE) + # Don't drop the year before 2017
  labs(title="The Dot Plot of Built Year by Registration Year Cateogory from 1805 to 2020 ", x= "Registration Year Cateogory", y="Built Year")
print(answer1.2_better)
```

**Q2 How are the number of storeys distributed among buildings?**

**Preparation**
```{r}
#overview
class(apt_buildings$no_of_storeys)
class(apt_buildings$`non-smoking_building`)
summary(apt_buildings$no_of_storeys)
unique(apt_buildings$`non-smoking_building`)

#clean
Q2cleaned <- apt_buildings %>%
  select(id, no_of_storeys, `non-smoking_building`) %>%
  filter(!is.na(no_of_storeys)&!is.na(`non-smoking_building`))
head(Q2cleaned)
```

**Q2.1 Summarizing**

*Task*: Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data (Task 1).
two other summary statistics: standard deviation, median
one numerical variable: no_of_storeys
one categorical variable: `non-smoking_building`

*Comment*: This is helpful. It highly links to Question 2. Computing mean and median supports measuring central tendency and computing range and standard deviation supports measuring statistical dispersion.
In addition attaching _`non-smoking_building`_ to the initial question can help with making assumptions of the relationship between building's smoking type and the number of storeys.

```{r}
#calculate within categories
answer2.1_seperate <- Q2cleaned %>%
  group_by(`non-smoking_building`)%>%
  summarize(
    Range=paste(min(no_of_storeys),"-", max(no_of_storeys)),
    Mean= mean(no_of_storeys),
    Stdev=sd(no_of_storeys),
    Median=median(no_of_storeys)
  )
#calculate all
answer2.1_total <-Q2cleaned %>%
  summarize(
    `non-smoking_building`= "TOTAL",
    Range=paste(min(no_of_storeys),"-", max(no_of_storeys)),
    Mean= mean(no_of_storeys),
    Stdev=sd(no_of_storeys),
    Median=median(no_of_storeys)
  )
#integrate 
answer2.1 <- bind_rows(answer2.1_seperate, answer2.1_total)
  
print(answer2.1)
```

**Q2.2 Graphing**

*Task*:Create a graph that has at least two geom layers (Task 8).

*Comment*: This is helpful. I used geom_boxplot to present results from summarizing task. In addition, I used geom_jitter. Adding individual data points on top of boxplots shows the distribution of data points within each category. This can be helpful to visualize the spread of the data.
```{r}
#offering an overall view (not distinguish whether buildings allow smoking or not)
answer2.2_total <- Q2cleaned %>%
  reframe(
    `non-smoking_building` = "TOTAL",
    no_of_storeys = no_of_storeys
  )
#visualize
answer2.2 <- bind_rows(Q2cleaned, answer2.2_total) %>%
  ggplot(aes(`non-smoking_building`, no_of_storeys))+
  geom_boxplot()+ # geom layer 1, visualizing summarizing task results
  geom_jitter(aes(color = `non-smoking_building`), alpha=0.04) + # geom layer 2
  labs(
    title= "The Box Plot of the Number of Storeys Categorized by Building Smoking Type",
    x= "Is the Building Non-Smoking",
    y= "The Number of Storeys"
  )+
  scale_color_manual(values = c("TOTAL" = "blue", "YES"="red", "NO" = "red"))
  
print(answer2.2)
```

**Q3 How is the distribution of separate gas, hydro, and water meter services among buildings?**

**Preparation**
```{r}
#clean
Q3cleaned <- apt_buildings %>%
  select(id, separate_gas_meters, separate_hydro_meters, separate_water_meters) %>%
  filter(!is.na(separate_gas_meters)&!is.na(separate_hydro_meters)&!is.na(separate_water_meters))
head(Q3cleaned)
```

**Q3.1 Summarizing**

*Task*: Compute the number of observations for at least one of your categorical variables. Do not use the function table() (Task 2). 
Categorical variables: separate_gas_meters, separate_hydro_meters, separate_water_meters

*Comment*: This is helpful. It offers an overview to understand the distribution of separate gas, hydro, and water meter services among buildings.
```{r}
library(tidyverse)
#gas
gas_count <- Q3cleaned %>%
  group_by(separate_gas_meters) %>%
  count()%>%
  pivot_wider(names_from = "separate_gas_meters", values_from = "n")
#hydro
hydro_count <- Q3cleaned %>%
  group_by(separate_hydro_meters) %>%
  count()%>%
  pivot_wider(names_from = "separate_hydro_meters", values_from = "n")
#water
water_count <- Q3cleaned %>%
  group_by(separate_water_meters) %>%
  count()%>%
  pivot_wider(names_from = "separate_water_meters", values_from = "n")
#combine
answer3.1 <- bind_rows(gas_count, hydro_count, water_count) %>%
  mutate(Service_type=c("gas", "hydro", "water"))%>%
  select(Service_type, everything())

print(answer3.1)
```

**Q3.2 Graphing**

*Task*:Create a graph that has at least two geom layers (Task 8).

*Comment*:This is helpful. This visualization supports comparing the availability of separate gas, hydro, water services. Adding a geom_text layer makes the exact number clear to the audience.
```{r}
#make tibble suitable for visualization
longer<-answer3.1%>% #using the outcome from summarizing task
  pivot_longer(cols=c(-Service_type),
             names_to='YES_NO',
             values_to='Count') 
print(longer)

answer3.2 <- ggplot(longer, aes(x = Service_type, y = Count, fill = YES_NO)) +
  geom_bar(stat = "identity") + #geom layer1
  labs(
    title = "Count of YES and NO for Each Service Type",
    x = "Service Type",
    y = "Count"
  )+
  geom_text(aes(label=Count),position = position_stack(vjust = 0.5), size=3) #geom layer 2

print(answer3.2)

```

**Q4What is the relationship between window types and year built?**

**Preparation**
```{r}
#overview:
unique(apt_buildings$window_type)
#clean
Q4cleaned<-apt_buildings %>%
  select(id, window_type, year_built)%>%
filter(!is.na(window_type)&!is.na(year_built))
head(Q4cleaned)
```
**Q4.1 Summarizing**

*Task*:Compute the number of observations for at least one of your categorical variables. Do not use the function table() (Task 2).
categorical variable:window_type

*Comment*: This is helpful. _window_type_ is a categorical variable, making sense to the task. This task does not directly answer Question 4. However, it brings insight to understand the variable window_type related to Question 4.
```{r}
answer4.1 <- Q4cleaned %>%
  group_by(window_type) %>%
  count()
print(answer4.1)
```

**Q4.2 Graphing**

*Task*: Make a graph where it makes sense to customize the alpha transparency (Task 7).

*Comment*: This is helpful. Because there are overlaps, the transparency makes sense.The darker color presents higher density. This graph can give audiences an overview and help them come up an assumption for Question 4.
```{r}
answer4.2 <- Q4cleaned%>%
  ggplot(aes(window_type,year_built))+
  geom_jitter(aes(color = window_type), alpha=0.2)+ #customize the transparency
  labs(title="Scatter Plot of Built Year by Window Type",
      x="Window Type",
      y="Built Year")
print(answer4.2)
```

<!----------------------------------------------------------------------------->

### 1.3 (2 points)

Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?

<!------------------------- Write your answer here ---------------------------->
While getting closer to the research questions, there are still aspects remain unclear and some needs to redefine. Particularly: 
* Q1: This question needs redefine. The key problem is that the range and variety of year_registered is not as wide as year_built. While year_built ranges from 1805-2019, buildings mainly registered in 2017. Digging into more details about the density of building registered year still has value, but redefining the question as _what is the relationship between year_registered and year_built?_ makes more sense to explore the City of Toronto's advancement in managing apartment buildings.
*Q2: This question is basically answered. Further exploration may include exploring the relationship between non-smoking_building and the number of storeys.
*Q3: This question is basically answered. Further exploration may include exploring how different combinations of separate_gas_meters,separate_hydro_meters,separate_water_meters services are distributed among buildings.
*Q4: This question remains unclear. The gap of the number of buildings categorized by window types is large. Need to find a suitable model to conduct statistical test to further answer this question.
<!----------------------------------------------------------------------------->

# Task 2: Tidy your data 

In this task, we will do several exercises to reshape our data. The goal here is to understand how to do this reshaping with the `tidyr` package.

A reminder of the definition of *tidy* data:

-   Each row is an **observation**
-   Each column is a **variable**
-   Each cell is a **value**

### 2.1 (2 points)

Based on the definition above, can you identify if your data is tidy or untidy? Go through all your columns, or if you have \>8 variables, just pick 8, and explain whether the data is untidy or tidy.

<!--------------------------- Start your work below --------------------------->

I choose the data set _Q1cleaned_ in Task 1 for *Q1 What is the density of building registered year?*  as the context to answer this question. 

It is tidy because it meets the three requirements. Each row represents a distinct and individual unit of data. Each column represents a distinct variable that is measured for each of the observations. Each cell holds a single, meaningful data point related to the observation and variable it intersects. 
<!----------------------------------------------------------------------------->

### 2.2 (4 points)

Now, if your data is tidy, untidy it! Then, tidy it back to it's original state.

If your data is untidy, then tidy it! Then, untidy it back to it's original state.

Be sure to explain your reasoning for this task. Show us the "before" and "after".

<!--------------------------- Start your work below --------------------------->
Before:
```{r}
head(Q1cleaned)
```
After:
```{r}
#untidy my data
untidied <- Q1cleaned %>%
  #make 2 rows present the same observation
  add_row(id=10359, year_built=1967, year_registered=2017)%>% 
  arrange(id)%>%
  #violate Each column is a **variable**
  tidyr::unite(col=BuiltYear_RegisteredYear,c("year_built","year_registered"), sep="_")
#violate each cell is a **value**
untidied[3,"BuiltYear_RegisteredYear"]<-"trhivhsk"
head(untidied)
```

<!----------------------------------------------------------------------------->

### 2.3 (4 points)

Now, you should be more familiar with your data, and also have made progress in answering your research questions. Based on your interest, and your analyses, pick 2 of the 4 research questions to continue your analysis in the remaining tasks:

<!-------------------------- Start your work below ---------------------------->
1.  *How is the distribution of separate gas, hydro, and water meter services among buildings?*
2.  *What is the relationship between window types and year built?*
<!----------------------------------------------------------------------------->

Explain your decision for choosing the above two research questions.

<!--------------------------- Start your work below --------------------------->
**As explained previously, Q1 needs to be redefined. Besides, 4~8 functions are required for the following task, to make a different, Q2 is not complicated enough. **
<!----------------------------------------------------------------------------->

Now, try to choose a version of your data that you think will be appropriate to answer these 2 questions. Use between 4 and 8 functions that we've covered so far (i.e. by filtering, cleaning, tidying, dropping irrelevant columns, etc.).

(If it makes more sense, then you can make/pick two versions of your data, one for each research question.) 

<!--------------------------- Start your work below --------------------------->
Before:
```{r}
head(Q3cleaned)
head(Q4cleaned)
```
After:
```{r}
# For Q3
Q3new<-Q3cleaned%>%
  #Function 1:mutate()
  mutate(separate_gas_meters = ifelse(separate_gas_meters == "YES", 1, 0),
         separate_hydro_meters = ifelse(separate_hydro_meters == "YES", 1, 0),
         separate_water_meters = ifelse(separate_water_meters == "YES", 1, 0)
         )%>%
  #Function 2: tidyr::unite()
  tidyr::unite(col=gas_hydro_water,c("separate_gas_meters","separate_hydro_meters", "separate_water_meters"), sep="")
print(Q3new)

#For Q4
Q4new <- Q4cleaned %>%
  #Function 3: filter()>>> because the window types are not mutually exclusive, DOUBLE PANE can also be THERMAL
  filter(window_type!="THERMAL")%>%
  mutate(window_type=ifelse(window_type=="DOUBLE PANE", 2, 1))%>%
  #Function 4: rename()
  rename(pane_type = window_type)%>%
  #Function 5: as.integer()
  mutate(pane_type = as.integer(pane_type))
print(Q4new) 

#Note: functions like select() already used to prepare data for Task 1. Because I am not sure if all functions I listed above counts, I practiced Function 6: full_join() to make one data set for Q3 and Q4.

#Function 6: full_join()
Q3Q4<-full_join(Q3new, Q4new, by="id")
print(Q3Q4)
```

<!----------------------------------------------------------------------------->

# Task 3: Modelling

## 3.0 (no points)

Pick a research question from 1.2, and pick a variable of interest (we'll call it "Y") that's relevant to the research question. Indicate these.

<!-------------------------- Start your work below ---------------------------->

**Research Question**: If a building was built in year 2020, what year is it most likely to be registered?

**Variable of interest**: year_registered

<!----------------------------------------------------------------------------->

## 3.1 (3 points)

Fit a model or run a hypothesis test that provides insight on this variable with respect to the research question. Store the model object as a variable, and print its output to screen. We'll omit having to justify your choice, because we don't expect you to know about model specifics in STAT 545.

-   **Note**: It's OK if you don't know how these models/tests work. Here are some examples of things you can do here, but the sky's the limit.

    -   You could fit a model that makes predictions on Y using another variable, by using the `lm()` function.
    -   You could test whether the mean of Y equals 0 using `t.test()`, or maybe the mean across two groups are different using `t.test()`, or maybe the mean across multiple groups are different using `anova()` (you may have to pivot your data for the latter two).
    -   You could use `lm()` to test for significance of regression coefficients.

<!-------------------------- Start your work below ---------------------------->

```{r}

```

<!----------------------------------------------------------------------------->

## 3.2 (3 points)

Produce something relevant from your fitted model: either predictions on Y, or a single value like a regression coefficient or a p-value.

-   Be sure to indicate in writing what you chose to produce.
-   Your code should either output a tibble (in which case you should indicate the column that contains the thing you're looking for), or the thing you're looking for itself.
-   Obtain your results using the `broom` package if possible. If your model is not compatible with the broom function you're needing, then you can obtain your results by some other means, but first indicate which broom function is not compatible.

<!-------------------------- Start your work below ---------------------------->

<!----------------------------------------------------------------------------->

# Task 4: Reading and writing data

Get set up for this exercise by making a folder called `output` in the top level of your project folder / repository. You'll be saving things there.

## 4.1 (3 points)

Take a summary table that you made from Task 1, and write it as a csv file in your `output` folder. Use the `here::here()` function.

-   **Robustness criteria**: You should be able to move your Mini Project repository / project folder to some other location on your computer, or move this very Rmd file to another location within your project repository / folder, and your code should still work.
-   **Reproducibility criteria**: You should be able to delete the csv file, and remake it simply by knitting this Rmd file.

<!-------------------------- Start your work below ---------------------------->

<!----------------------------------------------------------------------------->

## 4.2 (3 points)

Write your model object from Task 3 to an R binary file (an RDS), and load it again. Be sure to save the binary file in your `output` folder. Use the functions `saveRDS()` and `readRDS()`.

-   The same robustness and reproducibility criteria as in 4.1 apply here.

<!-------------------------- Start your work below ---------------------------->

<!----------------------------------------------------------------------------->

# Overall Reproducibility/Cleanliness/Coherence Checklist 

Here are the criteria we're looking for.

## Coherence (0.5 points)

The document should read sensibly from top to bottom, with no major continuity errors. 

The README file should still satisfy the criteria from the last milestone, i.e. it has been updated to match the changes to the repository made in this milestone. 

## File and folder structure (1 points)

You should have at least three folders in the top level of your repository: one for each milestone, and one output folder. If there are any other folders, these are explained in the main README.

Each milestone document is contained in its respective folder, and nowhere else.

Every level-1 folder (that is, the ones stored in the top level, like "Milestone1" and "output") has a `README` file, explaining in a sentence or two what is in the folder, in plain language (it's enough to say something like "This folder contains the source for Milestone 1").

## Output (1 point)

All output is recent and relevant:

-   All Rmd files have been `knit`ted to their output md files. 
-   All knitted md files are viewable without errors on Github. Examples of errors: Missing plots, "Sorry about that, but we can't show files that are this big right now" messages, error messages from broken R code
-   All of these output files are up-to-date -- that is, they haven't fallen behind after the source (Rmd) files have been updated.
-   There should be no relic output files. For example, if you were knitting an Rmd to html, but then changed the output to be only a markdown file, then the html file is a relic and should be deleted.

Our recommendation: delete all output files, and re-knit each milestone's Rmd file, so that everything is up to date and relevant.

## Tagged release (0.5 point)

You've tagged a release for Milestone 2. 

### Attribution

Thanks to Victor Yuan for mostly putting this together.
